---
title: "Airline hubs"
author: "Giulio Barcaroli"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Airline hubs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---


```{r setup, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(fig.width=8, fig.height=6,
  collapse = TRUE,
  comment = "#>"
)
# if (!require(bookdown)) install.packages("bookdown", dependencies=TRUE)
# library(bookdown)
library(QGA)
```

# Airline hubs

In this example, we show how, With the Quantum Genetic Algorithm, it is possible to determine the most convenient hubs, given a set of 25 USA airports.

## Prepare data 

As input data, we consider the 25 airports, with their geographical coordinates:

```{r}
cities <- read.csv("Airlines_cities.csv")
print(cities)
```

In addition, we get two square matrices. The first one contains the costs:

```{r}
cost <- read.csv("Airlines_cost.csv",header=F)
str(cost)
```

that is, for each origin airport, the unit cost (the price of the ticket), necessary to reach the destination airport.

The second one is related to the flows:

```{r}
flow <- read.csv("Airlines_flow.csv",header=F)
str(flow)
```

that is, for each origin airport, the estimated number of passengers that want to reach the destination airport.

## Optimization

We want to minimize the total cost of the optimal solution, given by the product of the flows muliplied by the unit cost, to determine which could be the three more convenient hubs for the 25 airports.

We define the fitness function consequently:

```{r}
airline_hub <- function(solution,input) {
  cost <- input$cost
  flow <- input$flow
  penalization <- sum(cost)*0.0005
  obj <- -sum(apply(cost[,solution] * flow[,solution], 1, min)) / sum(cost)
  if (length(table(solution)) < length(solution)) {
    obj <- obj - penalization
  }
  # cat("\nSolution:",solution,"  obj: ",obj)
  return(obj)
}
```

Some remarks:

1. this function will receive, as input, the cost matrix and the flow matrix; 

2. to evaluate the current solution, in both matrices (cost and flow) we select the three columns corresponding to the three hubs, and calculate the sum of the total costs for each origin airport to reach the candidate hub, selecting the minimum (the minus sign is due to the fact that we want to minimize the value of the objective function);

3. a penalization is introduced to eliminate the solutions with a number of hubs lower than three.

We can now proceed with the optimization. Note that we set the value of the genome equal to 3 (the number of the desired hubs), and the number of the values to be attributed to each element of the genome is set equal to the number of the airports (25).

```{r}
input <- list(cost=cost,flow=flow)
popsize = 20
Genome = 3 # Number of desired hubs
nvalues_sol = 25
set.seed(4321)
solutionQGA <- QGA(
                popsize,
                generation_max = 200,
                nvalues_sol,
                Genome,
                thetainit = 3.1415926535 * 0.025,
                thetaend = 3.1415926535 * 0.005,
                pop_mutation_rate_init = 3/(popsize + 1),
                pop_mutation_rate_end = 1/(popsize + 1),
                mutation_rate_init = 3/(Genome + 1),
                mutation_rate_end = 1/(Genome + 1),
                mutation_flag = TRUE,
                plotting = FALSE,
                progress = FALSE,
                verbose = FALSE,
                eval_fitness = airline_hub,
                eval_func_inputs = input,
                stop_iters = 150)
```


## Analysis of the solution

The found solution:

```{r}
solution <- solutionQGA[[1]]
cities$City[solution]
sol <- cost[,solution] 
find_min_column <- function(row) {
  which.min(row)
}
min_column_vector <- apply(sol, 1, find_min_column)
print(min_column_vector)
cities$Hub <- cities$City[solution][min_column_vector]
print(cities[,c(2,5)])
```

We can plot them on a map:

```{r message=FALSE, warning=FALSE}
if (!require(tidyverse)) install.packages("tidyverse", dependencies=TRUE)
if (!require(ggplot2)) install.packages("ggplot2", dependencies=TRUE)
if (!require(maps)) install.packages("maps", dependencies=TRUE)
library(tidyverse)
library(ggplot2)
library(maps)
connections <- cities %>%
  left_join(cities %>% select(City, Lat, Long), by = c("Hub" = "City")) %>%
  rename(Hub_Lat = Lat.y, Hub_Long = Long.y) %>%
  select(City, Lat = Lat.x, Long = Long.x, Hub, Hub_Lat, Hub_Long)
hubs <- unique(cities$Hub)
hub_connections <- expand.grid(Hub1 = hubs, Hub2 = hubs, stringsAsFactors = FALSE) %>%
  filter(Hub1 < Hub2) %>%  # Evita duplicati e auto-connessioni
  left_join(cities %>% select(City, Lat, Long), by = c("Hub1" = "City")) %>%
  rename(Lat1 = Lat, Long1 = Long) %>%
  left_join(cities %>% select(City, Lat, Long), by = c("Hub2" = "City")) %>%
  rename(Lat2 = Lat, Long2 = Long)
usa_map <- map_data("state")
gg <- ggplot() +
  geom_polygon(data = usa_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  coord_fixed(1.3) +
  xlim(-125, -66) + ylim(25, 50)
gg <- gg + geom_segment(data = connections, 
                        aes(x = Long, y = Lat, xend = Hub_Long, yend = Hub_Lat),
                        color = "blue", alpha = 0.5, size = 1)
gg <- gg + geom_segment(data = hub_connections,
                        aes(x = Long1, y = Lat1, xend = Long2, yend = Lat2),
                        color = "red", alpha = 0.7, size = 1)
gg <- gg + geom_point(data = cities, aes(x = Long, y = Lat), color = "blue", size = 2)
gg <- gg + geom_point(data = cities[solution,], aes(x = Long, y = Lat), color = "red", size = 3)
gg <- gg + geom_text(data = cities, aes(x = Long, y = Lat, label = City), hjust = 1, vjust = 1, size = 3)
gg <- gg + ggtitle("Airlines hubs (cost minimization)") +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
print(gg)
```

